---
title: "API Call in R"
output: html_notebook
---

### US Patents Office 

```{r}
# install.packages("patentsview")
# this is a package someone created for calling a US Patent API 
library(patentsview)
```

The main thing with API calls is that now you're using a specific query based language for the platform, rather than just R. Therefore, the syntax will look different. For the US Patents API, this is the PatentsView Query Language. This uses a JSON-based syntax.

You can read more here: https://patentsview.org/apis/api-query-language

More about the R package:
- there is one main function (search_pv()) that sends requests to the API 
- there are two parameters in this function: query, which tells the API how you want to filter the patents data, and fields, which tells the function which fields you want to retrieve

## Looking what happens with the default value for fields

```{r}
#?qry_funs
# this tells you more about the functions.
qry_1 = qry_funs$gt(patent_year = 2010)
qry_1 
# qry funs are a list of functions that make it easy to work with PatentView. There are 14 of them.

b = search_pv(query = qry_1, fields = NULL)

# default for fields gives you patent_id, patent_number, and patent_title 
```

## Now trying different values for fields 

```{r}
search_pv(
  query = qry_funs$gt(patent_year = 2010),
  fields = c("patent_abstract",
             "inventor_first_name")
)

# you can do more fields or choose different ones, these are just an example 
```

```{r}
#?search_pv
# this gives you a vector of fields that you can retrieve 
a = search_pv(
  query = qry_funs$gt(patent_year = 2010),
  fields = get_fields(endpoint = "patents", groups = c("patents", "inventors")))
# add parameters all_pages = TRUE to read all patents, default is 25 

# can also use the get_fields function 
# you can see that we are getting a lot more information about each patent 
x = data.frame(a$data$patents)
x

# issue with list type 

df <- apply(x,2,as.character)
# changed the list type to a character type

write.csv(df,"/Users/katieforester/Desktop/example.csv", row.names = FALSE)
# can handle a character type, not a list type 
```

## These are just basics, let's look at an example 

This is from an example I found online of how someone (Chris Baker, R Blogger) used the patent API in order to see where patents were coming from geographically. Here's his example. 

```{r}
# this chunk is what we learned above 

# reminder that for any function you're using, you can always put a ? mark in front of it to search it's use and parameters. This will be useful when you want to use these functions with different parameters in the future

library(patentsview)

# Write a query:
query <- with_qfuns( # with_qfuns is basically just: with(qry_funs, ...)
  and(
    begins(cpc_subgroup_id = 'H04L63/02'),
    # looking at a specific subgroup of patents
    gte(patent_year = 2007)
    # patent year awarded 
  )
)
# Create a list of fields:
fields <- c(
  c("patent_number", "patent_year"),
  get_fields(endpoint = "patents", groups = c("assignees", "cpcs"))
)
# Send HTTP request to API's server:
pv_res <- search_pv(query = query, fields = fields, all_pages = TRUE)
# this actually queries the API to pull the patent data 
```


```{r}
library(leaflet)
library(htmltools)
library(dplyr)
library(tidyr)

data <-
  pv_res$data$patents %>%
  # calling the patent dataframe 
  # using dplyr pipe, so this is the data we want 
    unnest(assignees) %>%
  # essentially ungroup the assignees so we can get all the selected variables below 
    select(assignee_id, assignee_organization, patent_number,
           assignee_longitude, assignee_latitude) %>%
  # select these variables 
    group_by_at(vars(-matches("pat"))) %>%
  # looks at all matches of the function, so we're grouping by the selected variables except those starting with pat 
  # group_by these variables 
    mutate(num_pats = n()) %>%
  # create a new column called number of patents that counts the number 
    ungroup() %>%
    select(-patent_number) %>%
  # deselect the number of patents 
    distinct() %>%
  # extract distinct rows by number of variables 
    mutate(popup = paste0("<font color='Black'>",
                          htmlEscape(assignee_organization), "<br><br>Patents:",
                          num_pats, "</font>")) %>%
  # this is just setting color settings for the map graph 
    mutate_at(vars(matches("_l")), as.numeric) %>%
  # make numeric variables 
    filter(!is.na(assignee_id))
  # get rid of any missing values 

leaflet(data) %>%
  # this is a graphical function 
  addProviderTiles(providers$CartoDB.DarkMatterNoLabels) %>%
  addCircleMarkers(lng = ~assignee_longitude, lat = ~assignee_latitude,
                   popup = ~popup, ~sqrt(num_pats), color = "yellow")
```


### Crunchbase

Again, for crunchbase, another individual has already created a more user friendly package for R that lets you query the API. 

I had an issue getting an API key, so I couldn't query it. This website looks like it has some useful info though:

https://blog.revolutionanalytics.com/2015/02/rcrunchbase-an-api-interface-to-crunchbase.html
https://data.crunchbase.com/v3.1/docs/using-the-api


```{r}
install.packages("rcrunchbase")
```





