---
title: "Descriptive Stats"
output:
  html_document:
    df_print: paged
---
 
## Upload dataset 

```{r}
library(AmesHousing)
ames = make_ames()

ames 
```

```{r}
head(ames)
# gives first six observations 
# good for checking that all variables you want are in data set and for looking at type of variables 
```

```{r}
str(ames)
# gives you the structure of the data 
# Data types - how the data are stored in R
# Levels are categories
```


## Min and Max of data 

```{r}
# use $ and variable name
min(ames$Lot_Frontage)
max(ames$Lot_Frontage)
ames
```

## Range 

```{r}
range(ames$Lot_Area)
```

## Mean 

```{r}
mean(ames$Lot_Area)
```

## Median 

```{r}
median(ames$Lot_Area)
```

```{r}
## can also use quantile function
quantile(ames$Lot_Area, 0.5)
quantile(ames$Lot_Area, 0.25)
# can do with any quartiles 
```
 
## Interquartile Range 

```{r}
# difference between first and third quartiles
IQR(ames$Lot_Area)
```

## Standard Dev and Variance 

```{r}
sd(ames$Lot_Area)
var(ames$Lot_Area)
```

```{r}
## can also apply to multiple columns 

# can also try : ames[,c(“lot_area”,”lot_frontage")


lapply(ames[, 3:4], sd)
# remember that lapply function applies a specific function to data --> i.e. (data you want, function you want)
```

## Really helpful summary function

```{r}
summary(ames[, 3:4])
```

## Mode 

Unfortunately, there is no function in R to find the mode of a variable (that I know of). Here's one way I know of to do it.

```{r}
table_ames <- table(ames$Lot_Area) # number of occurrences for each unique value
sort(table_ames, decreasing = TRUE) # sort highest to lowest

# although the printout is long, we can see that the mode for lot area is 9600
```

## Counts 

```{r}
summary(ames$Bldg_Type)
# gives you the number of each factor in the data 
```

## Conditional Counts for Nominal/Ordinal Variables

```{r}
# lets say we want to count the number of each "paved"
summary(ames$Paved_Drive)

# another way to count the number of paved 
sum(ames$Paved_Drive == "Paved")

# or condition 
sum(ames$Paved_Drive == "Paved" | ames$Paved_Drive == "Dirt_Gravel")

# and condition 

sum(ames$Paved_Drive == "Paved" & ames$Paved_Drive == "Dirt_Gravel")
sum(ames$Paved_Drive == "Paved" & ames$Alley == "No_Alley_Access")

```

## Conditional Counts for Numeric Variables 

```{r}
# find the number of lots greater than 1000
sum(ames$Lot_Area > 1000, na.rm=TRUE)
# na.rm = TRUE removes missing values for you 
```

```{r}
# find the number of lots between 1000 and 2000
sum(ames$Lot_Area > 1000 & ames$Lot_Area < 2000, na.rm=TRUE)
```

```{r}
library(dplyr)
ames2 = ames %>% mutate(filtered_area = ifelse(Lot_Area > 10000 & Lot_Area < 20000, 1, 0))%>%select(filtered_area, everything())

ames2
# now we have a variable at the end of our data set with this condition present 
```
```{r}
library(dplyr)
ames2 = ames %>% mutate(filtered_area = ifelse(Lot_Area > 10000 & Lot_Area < 20000, 1, 0))%>%filter(filtered_area ==1)
ames2
# now we have a variable at the end of our data set with this condition present
```

## Barplot 

```{r}
barplot(table(ames$Lot_Shape)) # table() is mandatory
# really basic automatic barplot in R 
```

```{r}
# more fancy, customizable barplot in R
library(ggplot2)

ggplot(ames, aes(x = Lot_Shape, fill = Lot_Shape)) +
  geom_bar() +
  ggtitle("Lot Shape of Houses in Ames") +
  xlab("Lot Shape") + ylab("Number of Houses")
  
```

## Line Plot 

```{r}
plot(ames$Lot_Area,
  type = "l"
) # "l" for line
```

## Histogram 

```{r}
hist(ames$Lot_Frontage)
```

```{r}
## can do same in ggplot 
ggplot(ames) +
  aes(x = Lot_Frontage) +
  geom_histogram(bins = 15)
# can change number of bins 
```

## Boxplot 

```{r}
boxplot(ames$Lot_Frontage)
```

```{r}
# side by side comparison of numerical to categorical 
boxplot(ames$Lot_Frontage ~ ames$Alley)
```

```{r}
ggplot(ames) +
  aes(x = Alley, y = Lot_Frontage) +
  geom_boxplot()
```

## Scatterplot 


```{r}
ames
plot(ames$Lot_Area, ames$Gr_Liv_Area)
```

```{r}
ggplot(ames) +
  aes(x = ames$Lot_Area, y = ames$Gr_Liv_Area) +
  geom_point()
```
```{r}
# add in a categorical factor for more info 
ggplot(ames) +
  aes(x = ames$Lot_Area, y = ames$Gr_Liv_Area, color = Alley) +
  geom_point() +
  scale_color_hue()
```

## QQplot for Normality Assumptions 

```{r}
# Draw points on the qq-plot:
qqnorm(ames$Lot_Area)
# Draw the reference line:
qqline(ames$Lot_Area)
```

We can see that this deviates a lot from normality (the theoretical line of normal data), so this normality assumption would be broken. 

## Density plots for distribution 

```{r}
plot(density(ames$Lot_Area))
```

